---
title: "P1_Simulartion_Exercise"
author: "Jerad Acosta"
date: "December 21, 2014"
output: pdf_document
---

Load Appropriate Libraries and set seed for reproducibility
```{r, message=FALSE}
library(ggplot2); library(dplyr)
set.seed(741)
```
### Creating 1000 Simulations:
To simulate 1000 exponential distributions with size 40 and **lambda** 0.2
we create a variable, **nosim**, to represent the number of simulartions,**n**
for the size of each distribution and set **lambda = 0.2**
```{r}
nosim <- 1000
n <- 40
lambda = 0.2
```

Next we create the simulated exponential distributions by filling a matrix
with dimensions 1000 by 40 where each of the 1000 rows represents a set of
40 variables (columns)
```{r}
simulation <- matrix(rexp(nosim * n, lambda), nosim)
print(paste(nrow(simulation),"rows","by", ncol(simulation),"columns"))
```
The Matrix was filled in with 40,000 variables of an exponential distribution
using the function **rexp(nosim * n, lambda)***
**NOTE:**
*That since each variable is iid the order in which we chose our groups of 
40 variables does not matter as long as it is not related to their values
This is the of taking 1000 samples of size 40 from the exponential distribution*

## Q1 with regard to the measured mean and the expected or theoretical mean
First we take the mean of each distribution as well as their cumulative mean
and variability
```{r}
colMeans <- apply(simulation, 1, mean)
meanAll <- mean(colMeans)
sdAll <- sd(colMeans)
```
```{r,echo=FALSE}
cat(paste("The observed mean of samples is", meanAll))
```
Where the theoretical mean is
$$\frac{1}{\lambda} = \frac{1}{`r lambda`} = `r 1/lambda`.$$

## Q2 compare variance of sample means with theoretical
```{r, echo=FALSE}
cat(paste("the observed standard deviation of means is", sdAll))
```
While the theoretical standard deviation is
$$\frac{\sigma}{\sqrt{n}} = \frac{1/\lambda}{\sqrt{`r n`}} = `r (1/lambda)/sqrt(n)`.$$
The Variance of our sample means is `r var(colMeans)`
With a theoretical mean of 
$$\frac{\sigma^{2}}{n} = \frac{1/\lambda^{2}}{`r n`} = `r (1/lambda^2)/n`.$$

## Q3 Show the distribution is approximately normal
Now we Use ggplot2 to plot the density of the sample means with the normal 
distribution $N\left(1/\lambda, \frac{1/\lambda}{\sqrt{`r n`}}\right)$ superimposed.
```{r}
# First we plot the relative density of each sample mean
ggplot(data = data.frame(colMeans), aes(x = colMeans)) + 
    geom_histogram(binwidth = 0.1, aes(y= ..density..), fill = "forest green", 
    color = "gray") + xlab("Sample Mean") +
# Place a Yellow vertical line to show where the distribution is centered
    geom_vline(xintercept = meanAll, size = 2, color = "red") +
# Use a Black Line to represent the theoretical mean
    geom_vline(xintercept = 5, size = 1, color = "black") +
# Superimpose the Normal Distibution
    stat_function(fun = dnorm, color = "gray9", size = 2, 
    arg = list(mean = 1/lambda, sd = 1/lambda*(1/sqrt(n)))) +
    ggtitle("Distribution of 1000 means from Sample size 40") 
```

It is quite apparent that both the **sample mean** or *center of the distribution*,
in **red** as well as the **variation** are well modeled by a normal distribution
which is outlined in **black** along with a black **theoretical** or *true*
**mean**


## But just as extra fun:
We can use the confidence interval for $1/\lambda$:
                $$\overline{x} \pm 1.96 \frac{s}{\sqrt{n}}$$
to see just how many of our samples captured the theoretical mean with their
appropriate confidence intervals with a T value of 1.96
```{r}
simulation <- data.frame(simulation)
sds <- apply(simulation, 1, sd)
simulation <- data.frame(simulation,colMeans, sds)
simulation <- simulation %>% mutate(CI = 
                (colMeans - 1.96 * sds / sqrt(n) < 1/lambda) & 
                (colMeans + 1.96 * sds / sqrt(n) > 1/lambda))
percent <- mean(simulation$CI)
```
#### According to these calculations `r percent*100`% of the samples means contained the true, theoretical mean